Trending assessment for nowcasts, measurements, and forecasts

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

The above measures use only the information on a point's quadrant and neglect further details on its location within the quadrant, as displayed in a four-quadrant plot.
However, points close to the zero point have intuitively less explanatory power.
If there is noise or non-systematic effects present in true value or nowcasts, those points are likely to be driven by noise to a large amount. \unsure{anstatt nowcasts eher predictions}
Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the above methods.
Points within that zone are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
The exclusion area's shape can be chosen according to the noise characteristics in predictions and true values.
In general, it is likely that the models of predictions have a noise component and should thus be part of the exclusion area.
We denote the measures accounting for an exclusion area by
\begin{align}
    \acceps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{\abs{\diffxt} > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\abs{\diffxt} > \varepsilon}}\label{eq:acceps}\\
    \accpeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > \varepsilon}} \label{eq:accpeps}\\
    \accmeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < \varepsilon }}\label{eq:accmeps}
\end{align}
The estimators can easily be adapted for other shapes of the exclusion area.
Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
A box-shaped exclusion area leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
Thus, only point 1 is excluded.
An exclusion area along one axis removes points in which one of the components could change the sign by a small amount of noise.
This particularly suits prediction models, where small amounts of noise are inevitable.
A cross-shaped exclusion area along both axes accounts for sign reversal in both components.
For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
\unsure{kann man vielleicht mu für verschiedene Varianten von Exclusion Area ausrechnen als Vergleich?}
In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point. \unsure{vielleicht genauer bei der Verwendung von Quantilen}
These shapes are theor etically appealing, but the interpretation of the resulting estimators gets ambiguous.
Thus, we use the simple exclusion areas above.

@@

\subsection{Conditional trending plot}\label{subsec:trending-cond-prob}
\unsure{Allgemeine Frage: können wir hier von Wahrscheinlichkeit sprechen? }
The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > \varepsilon)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < \varepsilon)$.
To assess the trending ability of a prediction, an evaluation on a finder grid is of interest.
Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries in the trending ability can be detected.

@@

>>>>

The above measures use only the information on a point's quadrant and neglect further details on its location within the quadrant, as displayed in a four-quadrant plot.
However, points close to the zero point have intuitively less explanatory power.
If there is noise or non-systematic effects present in true value or predictions, those points are likely to be driven by noise to a large amount. 
Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the above methods.
Points within that zone are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
The exclusion area's shape can be chosen according to the noise characteristics in predictions and true values.
In general, it is likely that the models of predictions have a noise component and should thus be part of the exclusion area.
We denote the measures accounting for an exclusion area by
\begin{align}
    \acceps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{\abs{\diffxt} > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\abs{\diffxt} > \varepsilon}}\label{eq:acceps}\\
    \accpeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > \varepsilon}} \label{eq:accpeps}\\
    \accmeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < \varepsilon }}\label{eq:accmeps}
\end{align}
The estimators can easily be adapted for other shapes of the exclusion area.
Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
A box-shaped exclusion area leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
Thus, only point 1 is excluded.
An exclusion area along one axis removes points in which one of the components could change the sign by a small amount of noise.
This particularly suits prediction models, where small amounts of noise are inevitable.
A cross-shaped exclusion area along both axes accounts for sign reversal in both components.
For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
\unsure{kann man vielleicht mu für verschiedene Varianten von Exclusion Area ausrechnen als Vergleich?}
In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point. \unsure{vielleicht genauer bei der Verwendung von Quantilen}
These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
Thus, we use the simple exclusion areas above.

@@

\subsection{Conditional trending plot}\label{subsec:trending-cond-prob}
\unsure{Allgemeine Frage: können wir hier von Wahrscheinlichkeit sprechen? }
The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > \varepsilon)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < \varepsilon)$.
To assess the trending ability of a prediction, an evaluation on a finder grid is of interest.
Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries in the trending ability can be detected.
\unsure{Asymmetrie vielleicht irgendwo oben als Begriff einführen}

@@

\unsure{berücsksichtigen die hier angeweneten Verfahren die Abhängigkeit?}
